Our Future Nuclear Data Needs

### 2018-12-16 nuclear_data_needs/nuclear_data_needs.tex
Finished through section 2.2

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[width=0.4\textwidth]{pipeline_cropped.pdf}
\caption{\label{fig:pipeline}The nuclear data pipeline.}
\end{wrapfigure}

@@

The nuclear reaction compilation database is referred to as Experimental Nuclear Reaction Data library (EXFOR) \cite{EXF}.  This includes not only reaction cross sections, but also related data such as fission yields, resonance integrals, polarization data and more.  Given the wide range of data types present in EXFOR there is significantly less error checking than what is performed during in XUNDL compilation.  However, an online visualization tool is provided to allow users to plot and manipulate the data. EXFOR is used by both nuclear reaction evaluators and the nuclear reaction research and application community.

@@

Elastic and inelastic scattering have a profound effect on neutron populations in any nuclear application by changing the directions and energies of the neutrons in the system.  However, improvements in neutron scattering data have hampered by the fact that they do not produce easily measured signals like fission, or a radiochemical observable like neutron capture.  The result has been significant disagreements between different reaction evaluations.  These disagreements manifest themselves as compensating errors that are resistant to the validation process.  This issue has been recognized by the nuclear data community since the early days of the WPEC \cite{Row89} and more recently by Maslov \cite{Mas11}.

One compelling example of compensating errors involves the Jezebel 239Pu critical assembly (PMF1).  Models of Jezebel using LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) reproduce a unity value of keff to better than 1 part in 104.  However, when the authors substituted one nuclear quantity one quantity at a time between the two databases, the value of keff varied considerably.  The greatest changes came from changing the elastic and inelastic scattering cross sections, which led to changes in the keff of -500 pcm and +704 pcm respectively. This is shown in Fig. 2 to the right.   Similar examples of such compensating errors were also seen in earlier work by Bauge \cite{Bau12}.

A new Collaborative International Evaluated Library Organization (CIELO) Coordinated Research Program was started under the auspices of the IAEA \cite{Cha18a} to have the international nuclear data evaluation community agree on a set of evaluated cross sections and output spectra for six critically-important nuclei: 1H, 16O, 56Fe, 235,238U and 239Pu in order to address the issue of compensating errors.  Recently, this list has been expanded to include 14,15N, 9Be, 23Na, 59Co, 58Ni, 238-242Pu.

CIELO has relatively little energy or angle differential data to draw upon to aid in its evaluation of (n,n’) on the “Big 3” Actinide, leading to “significant differences amongst the evaluations for 239Pu in the fast energy range, with ENDF/B-VII.1 and JENDL-4.0 lying significantly above the JEFF-3.1 evaluation.” \cite{Cha14}.  The 239Pu data referenced in their paper are all from indirect measurements made prior to 1970 \cite{Bat69, And61}.  The case is similar for 235U where they note that differences for En > 50 keV “have a large impact on the fast criticality (Godiva), leading to a 540 pcm swing in calculated criticality, as shown by Go Chiba \cite{Chi12}.”

One significant source of (n,xn) partial cross section data from 1997 to 2003 was the GEANIE collaboration (Germanium Array for Neutron-Induced Excitations) at the Los Alamos Neutron Science Center/Weapons Neutron Research (LANSCE/WNR) facility. While the primary goal of the GEANIE collaboration was the determination of the 239Pu(n,2n) cross section \cite{Ber02}, it resulted in the publication of more than two dozen papers over the course of more than 15 years.  Experiments were run on all three of the major Actinides: 235U \cite{You01}, 239Pu \cite{Ber02} and 238U \cite{Fot04}.  It should be noted that GEANIE did not include any neutron detectors, producing only angle-integrated partial -ray cross sections.

The need for new scattering data has motivated many new experiments.  This has included energy-differential/angle-integrated measurements of the 238U(n,n’) cross section using the GELINA spallation neutron source in Geel Belgium by Bacquias et al., \cite{Nem13} and quasi-angle-differential measurements using the Rensselaer Polytechnic Institute Linac by Daskalakis et al., \cite{Das14}.

The ideal way to improve (n,n’) data on the 235,238U and 239Pu would be to perform measurements of the scattered neutron in coincidence with transitions emitted by the excited nucleus.  In the case of the Actinides this will potentially require the observation of conversion electrons given that the lowest transitions in all actinides are almost entirely converted.  Unfortunately, neutron beams usually have quite low intensity, with a fluence in the fast (1-10 MeV) region between 103 and 107 n/s/cm2 at the target location.  Furthermore, many spallation neutron sources have a significant high energy spectral component that can cause inter-element scattering in a neutron scintillator array rendering precise determination of the scattered neutron energy hard to  perform. However, intense limited-energy neutron sources based on thick-target deuteron breakup \cite{Har18} and forward-fitting techniques developed for use at spallation neutron sources \cite{Kee18} off an opportunity to perform neutron-gamma coincident measurements in the fast fission energy spectrum.  These techniques could provide valuable data to improve nuclear data evaluations of inelastic scattering, even on targets where fission is an open channel.

@@

Information regarding states in the unresolved continuum region above the particle separation energy is of central importance to accurately modeling the competition between particle and gamma-ray emission.  The most important of these are the nuclear level density (NLD) as a function of energy, spin and parity, and the radiative strength function (RSF), which describes the ability of excited nuclear matter to absorb or emit photons.  A number of recent studies have shown the importance of the RSF to modeling (n,) reactions for neutrons with energies between 1 and 100 keV, which corresponds to the spectrum at the site of neutron capture nucleosynthesis in stars \cite{Lar10, Lar15} and energy generation in fast reactors and nuclear security applications.

Over the past several years’ work led by the group at the University of Oslo has shown that there is a surprisingly large amount of M1 radiative strength believed to be a type of “scissors” collective mode throughout the Actinide mass region near the valley of stability including 238U, 239,240,242Pu and 232Th \cite{Lap16, Gut13, Gut13a, Gut14, Gut15}.

@@

(n,n’) provides a useful tool for studying the spin cut-off parameter since the reactions proceed largely through the population of a compound nucleus with a relatively wide range of angular momentum. Figure 4 shows the ratio of the partial cross section for the population of the  level in 238U(n,n’) as a function of incident neutron energy calculated using the TALYS reaction code for two different choices of spin cut-off.  The solid line shows the case where a constant temperature level density is assumed and the spin cut-off parameter was varied from 75% to 125% of its “normal” value given by the equation (3) above.  The ratio decreases by nearly a factor of 2 with increasing energy.  In contrast, the dashed line shows the ratio of the  population of the  state when the LD and RSF from the recent Oslo measurements \cite{Gut13a} is divided by the results obtained using the default Gilbert and Cameron in TALYS.  In this case the ratio decreases by nearly a factor of 2.  Similar behavior is seen in the partial cross sections for other off-yrast levels as well.

Information regarding the J-dependence of the NLD can also be obtained through observation of the relative population of two low-lying states with very different spins via activation.  This method has been used in Hg and Au by Chakravarty \cite{Cha92} and more recently by Sudar in Hg and Pt \cite{Sud06}.  Figure 5 shows this for the case of 52mMn(Jπ=2+)/52gMn(Jπ=6+) ratio populated via 56Fe(p,n) measured via stacked target activation at the LBNL 88-Inch cyclotron as compared to TALYS calculations for different values of the spin cut-off parameter (left) and the 6 level density models available for use in TALYS on the right.  
It should be noted that both the (n,n’) and isomer-to-ground state ratio measurements could be performed as part of a program that supports improved neutron transport as described in Section II.B above, and a program of cross section measurements in support of medical isotope production described in sections IV below.  This reflects the interdependency between nuclear data over a wide range of applications.

@@

The antineutrino spectrum produced in a nuclear reactor is calculated as the sum of the spectra produced by each of the nuclear fuels, 235,238U and 239,241Pu, weighted by the respective fission fractions \cite{Vog81}.   The spectra for each fuel can be calculated using nuclear databases, in what is commonly known as the ‘summation method’.   Alternatively, for 235U and 239,241Pu the spectra can also be calculated by converting the aggregate electron spectra measured at ILL in the 1980s \cite{Fel82,Sch85,Hah89}.   The current best predictions for the 235U and 239,241Pu are those by Huber \cite{Hub11} using the conversion method, while for 238U is that by Mueller et al. \cite{Mue11} using a hybrid conversion-summation method.

@@

\subsection{Improved (n,x) data for active interrogation}

@@

Over 20 million nuclear medicine procedures are performed each year in the United States \cite{Met09}, with the vast majority involving imaging.  99mTc (t½= 6.0067(5) hour), which decays predominately via the emission of single 140 keV gamma-ray, is used in 80% of of these procedures. 99mTc is produced via 235U(n,f).  However, this approach is less than ideal due to the increasing age of many reactors \cite{Qai12} and a nuclear weapons proliferation risk associated with reactors utilizing Highly-Enriched Uranium (HEU).  As a result, alternative 99mTc production pathways, as well as studies of the associated production of long-lived impurities \cite{Upd13}, are under consideration \cite{Rut09}. Other imaging radionuclides include 18F (t½= 109.771(20) min) which is used in Positron Emission Tomography (PET) which is produced locally via the 18O(p,n) reaction, 82Rb (t½=1.2575(2) min) which is used in cardiac imaging and 68Ga  (t½= 67.71(8) min).  The short half-lives of 82Rb and 68Ga necessitate that they be extracted from a longer-lived radioactive parent nuclide generator whose longer half-lives of 270.93(13) and 25.35(3) days respectively, allow them to be produced at regional facilities and distributed to medical facilities throughout the world.  
In addition to imaging, a large number of therapeutic radiopharmaceuticals are in use or under development for cancer treatment.  The principle behind therapeutic radionuclide cancer therapy involves depositing a targeted dose, e.g., energy per unit mass, to a tumor cell that capable of producing damage to multiple strands of DNA, from which the cell cannot recover.  The ideal dose delivery involves radiation with a high Linear Energy Transfer (LET).  While some - nuclides are already in use, such as 177Lu (t½= 6.6475(20) days), the relatively long range of -particles makes them less favorable than other modes of radioactive decay.  These include Auger/Coster-Kronig electron-emitters which typically have a range of less than 1 µm, or -particles emitters which have ranges of 2-10 µm.   
There is a particularly large amount of interest in alpha-emitting nucleus 225Ac.  225Ac is particularly well-suited for medical applications in that it has a long-enough lifetime (t½= 10.0(1) days) to facilitate its incorporation into targeting molecules, and its daughters then decay relatively rapidly (t1/2<3.2 hours) to stable products, minimizing concurrent dose to healthy tissue. Other promising alpha-therapeutic radionuclides include 211At, 212Pb/212Bi, 213Bi, 226Th and 227Th have all been identified as high priority topics by the most recent Nuclear Science Advisory Committee’s Long Range Plan on Isotope Production \cite{Nsa15}.
A particularly promising approach nuclear medical treatment modality involves coupling a pair of chemically similar radioisotopes including of a therapeutic nuclide capable of delivering a high localized radiation dose and a chemically-identical isotope that emits a positron for PET or a gamma suitable for single photon emission computed tomography (SPECT) imaging. One such promising “theranostic” pair involves the PET isotope 134Ce (t½= 3.16(4) days) combined with 225Ac.  A recent NSAC report \cite{Nsa15} highlights additional pairs of isotopes that have potential uses as theranostic agents.  

@@

All of these regional sources utilize high-energy (≥100 MeV) (p,x) reactions on stable targets.  In addition to producing the radionuclides of interest these reactions create a large flux of “secondary” spallation protons and neutrons that can in turn initiate nuclear reactions providing either a useful alternative source of valuable radionuclide or a troublesome chemically-identical contaminant.  The secondary neutron flux is particularly important since the lack of electronic stopping for neutrons makes them advantageous for radionuclide production \cite{Voy18a}.  The energy spectrum of secondary neutrons from IPF was recently measured using spectral unfolding \cite{Mos16} and for thick target deuteron break-up on beryllium at the LBNL 88-Inch cyclotron \cite{Har18} via a secondary time-of-flight technique. A number of therapeutic radionuclides can potentially be produced via secondary neutron-induced (n,p) reactions. Examples of possible targets include 32S, 47Ti, natZn (for the production of 64.67Cu), 105Pd, 149Sm, 175Lu, and 177Hf.  In addition, the production of alpha-emitting radionuclides such as 225Ac, 223Ra, and 227Th using neutrons has yet to be explored. 
Unfortunately, accurate modeling of these high-energy reactions is notoriously difficult.  Figures 7 shows a comparison between cross sections for the production of 225Ac from (n,x) reactions on 232Th calculated with the reaction codes typically used to predict isotope production.  This includes EMPIRE-3.2.3, CoH 3.5.1, ALICE-2017, the TENDL-2011 and TENDL-2017 libraries produced using TALYS, as well as the most recent version of the code (TALYS 1.9) \cite{cite-talys} and the European Activation File (EAF-2010).  Differences on the order of 50-100% are often seen.  

@@

The flux is typically determined using a well-known monitor reaction on a co-irradiated foil that leads to the production of an easily quantified long-lived residual nucleus. However, for both charged-particle and neutron-induced reactions, there relatively few monitor reactions that can be applied in the low to intermediate energy regimes of 30-70 MeV.  The situation is worse above 70 MeV monitor reactions where values free from the potential influence of secondary neutron contributions are even harder to come by.  Knowledge of these cross sections as a function of the incident particle energy is central to the isotope production, with recent efforts focused on developing the 93Nb(p,4n)90Mo (t½= 5.56(9) hours) as a monitor reaction \cite{Voy18b, Kim18}.

@@

One artifact of the nuclear data tuning process, described in previous reference to the work by Bauge, is that confidence is reduced for use of data for applications other than the those for which the evaluations were tuned. Where fission and capture reactions are modified to ensure that the integral value of keff is maintained, little confidence remains in the independent use of the fission cross section itself, which is essential to the prediction of reactor power distributions. These power distributions are essential to the prediction of peak fuel temperatures during a reactor transient to ensure that the fuel does not exceed thermal limits that are set to ensure that fuel integrity is not compromised. Reduced confidence in the fission cross section can have real consequences in the allowed power level and operating regimes for these systems.

@@

Many of these advanced systems also require so called high-assay low enriched uranium (HA-LEU), which has a 235U enrichment >5% and <20% by mass, were the current fleet of LWRs utilize enrichments of 5% or less. With few benchmarks experiments in HA-LEU enrichment range, the interplay of the cross sections for multiple nuclei become important, especially for LEU experiments where 1H, 16O, 235U, 238U and others are applied simultaneously in the calculation of the integral keff response.  Similar to the Chadwick \cite{Cha18} and Bauge \cite{Bau12} studies of Jezebel critical assembly, a team at ORNL preformed an investigation of a hypothetical transportation package with 20% enriched 235U in the form of UF6 using ENDF/B-VII.1 and ENDF/B-VIII.0 libraries by only exchanging the CIELO nuclides. A swing in keff of 450 pcm is observed.  Future investigation and validation are required to gain confidence in either calculation.

@@

Following NDNCA, Dr. Catherine Romano, working with the DOE-NP and the NNSA office on counter-proliferation research and development (NA-22), organized a grassroots “Nuclear Data Working Group” (NDWG) comprised of subject matter experts selected by program managers from DOE-NP, DOE-Nuclear Energy, the Isotopes Program, the Department of Homeland Security and the several offices within the National Nuclear Security Administration to identify a “roadmap” to address the most important and crosscutting nuclear data needs including all steps in the nuclear data pipeline from measurement through compilation, evaluation and processing.  The NDWG presented their results to participating program managers as well as other interested agencies during the Nuclear Data Exchange Meeting (NDEM) in April 2016.  The scope of the recommended work included:
1. infrastructure modernization, 
2. expanded covariance data, 
3. inelastic scatter on 235U, 238U and 239Pu, 
4. gamma emission data, 
5. fission yields, and 
6. renewed nuclear data target production capabilities.
The NDEM provided a venue for valuable discussions between the nuclear data community and program managers and the tasks required to provide evaluated nuclear data to the database.

>>>>

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[width=0.39\textwidth]{pipeline_cropped.pdf}
\caption{\label{fig:pipeline}The nuclear data pipeline.}
\end{wrapfigure}

@@

The nuclear reaction compilation database is referred to as the Experimental Nuclear Reaction Data library (EXFOR) \cite{EXF}.  This includes not only reaction cross sections, but also related data such as fission yields, resonance integrals, polarization data and more.  Given the wide range of data types present in EXFOR there is significantly less error checking than what is performed during in XUNDL compilation.  However, an online visualization tool is provided to allow users to plot and manipulate the data. EXFOR is used by both nuclear reaction evaluators and the nuclear reaction research and application community.

@@

Elastic and inelastic scattering have a profound effect on neutron populations in any nuclear application by changing the directions and energies of the neutrons in the system.  However, improvements in neutron scattering data have been hampered by the fact that they do not produce easily measured signals like fission, or a radiochemical observable like neutron capture.  The result has been significant disagreements between different reaction evaluations.  These disagreements manifest themselves as compensating errors that are resistant to the validation process.  This issue has been recognized by the nuclear data community since the early days of the WPEC \cite{Row89} and more recently by Maslov \cite{Mas11}.

One compelling example of compensating errors involves the Jezebel \ce{^{239}Pu} critical assembly (PMF1).  Models of Jezebel using LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) reproduce a unity value of $k_{eff}$ to better than 1 part in 104.  However, when the authors substituted one nuclear quantity one quantity at a time between the two databases, the value of $k_{eff}$ varied considerably.  The greatest changes came from changing the elastic and inelastic scattering cross sections, which led to changes in the $k_{eff}$ of -500 pcm and +704 pcm, respectively. This is shown in \autoref{fig:rotated_swings}.   Similar examples of such compensating errors were also seen in earlier work by Bauge \cite{Bau12}.

\begin{figure}
 \centering
%  \includegraphics[clip=true,trim=1.5in 3.4in 1.8in 3.5in, scale=0.8]{./figures/variation_curve.pdf}
 \includegraphics[width=0.5\linewidth]{rotated_swings.pdf}

 % variation_curve.pdf: 612x792 pixel, 72dpi, 21.59x27.94 cm, bb=0 0 612 792
 \caption{Simulations of criticality $k_{eff}$ for \ce{^{239}Pu} for the fast assembly Jezebel, PMF-1. This figure shows that both LANL CIELO-1 (ENDF/B-VIII.0) and CEA CIELO-2 (JEFF-3.3) predict similar $k_{eff}$ values, but do so for very different reasons. The changes in criticality are evident when individual cross section channels are substituted between the two evaluations.  The shaded region shows the  resulting range of values for $k_{eff}$.}
 \label{fig:rotated_swings}
\end{figure}

A new Collaborative International Evaluated Library Organization (CIELO) Coordinated Research Program was started under the auspices of the IAEA \cite{Cha18a} to have the international nuclear data evaluation community agree on a set of evaluated cross sections and output spectra for six critically-important nuclei: \ce{^{1}H}, \ce{^{16}O}, \ce{^{56}Fe}, \ce{^{235,238}U} and \ce{^{239}Pu} in order to address the issue of compensating errors.  Recently, this list has been expanded to include \ce{^{14,15}N}, \ce{^{9}Be}, \ce{^{23}Na}, \ce{^{59}Co}, \ce{^{58}Ni}, \ce{^{238-242}Pu}.

CIELO has relatively little energy or angle differential data to draw upon to aid in its evaluation of (n,n') on the \enquote{Big 3} Actinides, leading to \enquote{significant differences amongst the evaluations for \ce{^{239}Pu} in the fast energy range, with ENDF/B-VII.1 and JENDL-4.0 lying significantly above the JEFF-3.1 evaluation.} \cite{Cha14}.  The \ce{^{239}Pu} data referenced in their paper are all from indirect measurements made prior to 1970 \cite{Bat69, And61}.  The case is similar for \ce{^{235}U} where they note that differences for $E_n >$ 50 keV \enquote{have a large impact on the fast criticality (Godiva), leading to a 540 pcm swing in calculated criticality, as shown by Go Chiba \cite{Chi12}.}

One significant source of (n,xn$\gamma$) partial cross section data from 1997 to 2003 was the GEANIE collaboration (Germanium Array for Neutron-Induced Excitations) at the Los Alamos Neutron Science Center/Weapons Neutron Research (LANSCE/WNR) facility. While the primary goal of the GEANIE collaboration was the determination of the \ce{^{239}Pu}(n,2n) cross section \cite{Ber02}, it resulted in the publication of more than two dozen papers over the course of more than 15 years.  Experiments were run on all three of the major Actinides: \ce{^{235}U} \cite{You01}, \ce{^{239}Pu} \cite{Ber02} and \ce{^{238}U} \cite{Fot04}.  It should be noted that GEANIE did not include any neutron detectors, producing only angle-integrated partial $\gamma$-ray cross sections.

The need for new scattering data has motivated many new experiments.  This has included energy-differential/angle-integrated measurements of the \ce{^{238}U}(n,n') cross section using the GELINA spallation neutron source in Geel Belgium by Bacquias et al., \cite{Nem13} and quasi-angle-differential measurements using the Rensselaer Polytechnic Institute Linac by Daskalakis et al., \cite{Das14}.

The ideal way to improve (n,n') data on the \ce{^{235,238}U} and \ce{^{239}Pu} would be to perform measurements of the scattered neutron in coincidence with transitions emitted by the excited nucleus.  In the case of the Actinides this will potentially require the observation of conversion electrons given that the lowest transitions in all actinides are almost entirely converted.  Unfortunately, neutron beams usually have quite low intensity, with a fluence in the fast (1--10 MeV) region between $10^3$ and $10^7$ n/s/cm$^2$ at the target location.  Furthermore, many spallation neutron sources have a significant high-energy spectral component that can cause inter-element scattering in a neutron scintillator array rendering precise determination of the scattered neutron energy difficult to  perform. However, intense limited-energy neutron sources based on thick-target deuteron breakup \cite{Harrig2018} and forward-fitting techniques developed for use at spallation neutron sources \cite{Kee18} off an opportunity to perform neutron-gamma coincident measurements in the fast fission energy spectrum.  These techniques could provide valuable data to improve nuclear data evaluations of inelastic scattering, even on targets where fission is an open channel.

@@

Information regarding states in the unresolved continuum region above the particle separation energy is of central importance to accurately modeling the competition between particle and gamma-ray emission.  The most important of these are the nuclear level density (NLD) as a function of energy, spin and parity, and the radiative strength function (RSF), which describes the ability of excited nuclear matter to absorb or emit photons.  A number of recent studies have shown the importance of the RSF to modeling (n,) reactions for neutrons with energies between 1 and 100 keV, which corresponds to the spectrum at the site of neutron capture nucleosynthesis in stars \cite{PhysRevC.82.014318, Lar15} and energy generation in fast reactors and nuclear security applications.

Over the past several years' work led by the group at the University of Oslo has shown that there is a surprisingly large amount of M1 radiative strength believed to be a type of \enquote{scissors} collective mode throughout the Actinide mass region near the valley of stability including 238U, 239,240,242Pu and 232Th \cite{Lap16, Gut13, Gut13a, Gut14, Gut15}.

@@

(n,n') provides a useful tool for studying the spin cut-off parameter since the reactions proceed largely through the population of a compound nucleus with a relatively wide range of angular momentum. Figure 4 shows the ratio of the partial cross section for the population of the  level in 238U(n,n') as a function of incident neutron energy calculated using the TALYS reaction code for two different choices of spin cut-off.  The solid line shows the case where a constant temperature level density is assumed and the spin cut-off parameter was varied from 75% to 125% of its \enquote{normal} value given by the equation (3) above.  The ratio decreases by nearly a factor of 2 with increasing energy.  In contrast, the dashed line shows the ratio of the  population of the  state when the LD and RSF from the recent Oslo measurements \cite{Gut13a} is divided by the results obtained using the default Gilbert and Cameron in TALYS.  In this case the ratio decreases by nearly a factor of 2.  Similar behavior is seen in the partial cross sections for other off-yrast levels as well.

Information regarding the J-dependence of the NLD can also be obtained through observation of the relative population of two low-lying states with very different spins via activation.  This method has been used in Hg and Au by Chakravarty \cite{Cha92} and more recently by Sudar in Hg and Pt \cite{Sud06}.  Figure 5 shows this for the case of 52mMn(Jπ=2+)/52gMn(Jπ=6+) ratio populated via 56Fe(p,n) measured via stacked target activation at the LBNL 88-Inch cyclotron as compared to TALYS calculations for different values of the spin cut-off parameter (left) and the 6 level density models available for use in TALYS on the right.  
It should be noted that both the (n,n') and isomer-to-ground state ratio measurements could be performed as part of a program that supports improved neutron transport as described in Section II.B above, and a program of cross section measurements in support of medical isotope production described in sections IV below.  This reflects the interdependency between nuclear data over a wide range of applications.

@@

The antineutrino spectrum produced in a nuclear reactor is calculated as the sum of the spectra produced by each of the nuclear fuels, 235,238U and 239,241Pu, weighted by the respective fission fractions \cite{Vog81}.   The spectra for each fuel can be calculated using nuclear databases, in what is commonly known as the ‘summation method'.   Alternatively, for 235U and 239,241Pu the spectra can also be calculated by converting the aggregate electron spectra measured at ILL in the 1980s \cite{Fel82,Sch85,Hah89}.   The current best predictions for the 235U and 239,241Pu are those by Huber \cite{Hub11} using the conversion method, while for 238U is that by Mueller et al. \cite{Mue11} using a hybrid conversion-summation method.

@@

\subsection{Improved (n,x$\gamma$) data for active interrogation}

@@

Over 20 million nuclear medicine procedures are performed each year in the United States \cite{Met09}, with the vast majority involving imaging.  99mTc (t$_{1/2}$= 6.0067(5) hour), which decays predominately via the emission of single 140 keV gamma-ray, is used in 80\% of of these procedures. 99mTc is produced via 235U(n,f).  However, this approach is less than ideal due to the increasing age of many reactors \cite{Qai12} and a nuclear weapons proliferation risk associated with reactors utilizing Highly-Enriched Uranium (HEU).  As a result, alternative 99mTc production pathways, as well as studies of the associated production of long-lived impurities \cite{Upd13}, are under consideration \cite{Rut09}. Other imaging radionuclides include 18F (t$_{1/2}$= 109.771(20) min) which is used in Positron Emission Tomography (PET) which is produced locally via the 18O(p,n) reaction, 82Rb (t$_{1/2}$=1.2575(2) min) which is used in cardiac imaging and 68Ga  (t$_{1/2}$= 67.71(8) min).  The short half-lives of 82Rb and 68Ga necessitate that they be extracted from a longer-lived radioactive parent nuclide generator whose longer half-lives of 270.93(13) and 25.35(3) days respectively, allow them to be produced at regional facilities and distributed to medical facilities throughout the world.

In addition to imaging, a large number of therapeutic radiopharmaceuticals are in use or under development for cancer treatment.  The principle behind therapeutic radionuclide cancer therapy involves depositing a targeted dose, e.g., energy per unit mass, to a tumor cell that capable of producing damage to multiple strands of DNA, from which the cell cannot recover.  The ideal dose delivery involves radiation with a high Linear Energy Transfer (LET).  While some - nuclides are already in use, such as 177Lu (t$_{1/2}$= 6.6475(20) days), the relatively long range of -particles makes them less favorable than other modes of radioactive decay.  These include Auger/Coster-Kronig electron-emitters which typically have a range of less than 1 µm, or -particles emitters which have ranges of 2-10 µm.

There is a particularly large amount of interest in alpha-emitting nucleus 225Ac.  225Ac is particularly well-suited for medical applications in that it has a long-enough lifetime (t$_{1/2}$= 10.0(1) days) to facilitate its incorporation into targeting molecules, and its daughters then decay relatively rapidly (t1/2<3.2 hours) to stable products, minimizing concurrent dose to healthy tissue. Other promising alpha-therapeutic radionuclides include 211At, 212Pb/212Bi, 213Bi, 226Th and 227Th have all been identified as high priority topics by the most recent Nuclear Science Advisory Committee's Long Range Plan on Isotope Production \cite{Nsa15}.

A particularly promising approach nuclear medical treatment modality involves coupling a pair of chemically similar radioisotopes including of a therapeutic nuclide capable of delivering a high localized radiation dose and a chemically-identical isotope that emits a positron for PET or a gamma suitable for single photon emission computed tomography (SPECT) imaging. One such promising \enquote{theranostic} pair involves the PET isotope 134Ce (t$_{1/2}$= 3.16(4) days) combined with 225Ac.  A recent NSAC report \cite{Nsa15} highlights additional pairs of isotopes that have potential uses as theranostic agents.

@@

All of these regional sources utilize high-energy (≥100 MeV) (p,x) reactions on stable targets.  In addition to producing the radionuclides of interest these reactions create a large flux of \enquote{secondary} spallation protons and neutrons that can in turn initiate nuclear reactions providing either a useful alternative source of valuable radionuclide or a troublesome chemically-identical contaminant.  The secondary neutron flux is particularly important since the lack of electronic stopping for neutrons makes them advantageous for radionuclide production \cite{Voy18a}.  The energy spectrum of secondary neutrons from IPF was recently measured using spectral unfolding \cite{Mos16} and for thick target deuteron break-up on beryllium at the LBNL 88-Inch cyclotron \cite{Har18} via a secondary time-of-flight technique. A number of therapeutic radionuclides can potentially be produced via secondary neutron-induced (n,p) reactions. Examples of possible targets include 32S, 47Ti, natZn (for the production of 64.67Cu), 105Pd, 149Sm, 175Lu, and 177Hf.  In addition, the production of alpha-emitting radionuclides such as 225Ac, 223Ra, and 227Th using neutrons has yet to be explored.

Unfortunately, accurate modeling of these high-energy reactions is notoriously difficult.  Figures 7 shows a comparison between cross sections for the production of 225Ac from (n,x) reactions on 232Th calculated with the reaction codes typically used to predict isotope production.  This includes EMPIRE-3.2.3, CoH 3.5.1, ALICE-2017, the TENDL-2011 and TENDL-2017 libraries produced using TALYS, as well as the most recent version of the code (TALYS 1.9) \cite{cite-talys} and the European Activation File (EAF-2010).  Differences on the order of 50--100\% are often seen.

@@

The flux is typically determined using a well-known monitor reaction on a co-irradiated foil that leads to the production of an easily quantified long-lived residual nucleus. However, for both charged-particle and neutron-induced reactions, there relatively few monitor reactions that can be applied in the low to intermediate energy regimes of 30-70 MeV.  The situation is worse above 70 MeV monitor reactions where values free from the potential influence of secondary neutron contributions are even harder to come by.  Knowledge of these cross sections as a function of the incident particle energy is central to the isotope production, with recent efforts focused on developing the 93Nb(p,4n)90Mo (t$_{1/2}$= 5.56(9) hours) as a monitor reaction \cite{Voy18b, Kim18}.

@@

One artifact of the nuclear data tuning process, described in previous reference to the work by Bauge, is that confidence is reduced for use of data for applications other than the those for which the evaluations were tuned. Where fission and capture reactions are modified to ensure that the integral value of $k_{eff}$ is maintained, little confidence remains in the independent use of the fission cross section itself, which is essential to the prediction of reactor power distributions. These power distributions are essential to the prediction of peak fuel temperatures during a reactor transient to ensure that the fuel does not exceed thermal limits that are set to ensure that fuel integrity is not compromised. Reduced confidence in the fission cross section can have real consequences in the allowed power level and operating regimes for these systems.

@@

Many of these advanced systems also require so called high-assay low enriched uranium (HA-LEU), which has a 235U enrichment >5% and <20% by mass, were the current fleet of LWRs utilize enrichments of 5% or less. With few benchmarks experiments in HA-LEU enrichment range, the interplay of the cross sections for multiple nuclei become important, especially for LEU experiments where 1H, 16O, 235U, 238U and others are applied simultaneously in the calculation of the integral $k_{eff}$ response.  Similar to the Chadwick \cite{Cha18} and Bauge \cite{Bau12} studies of Jezebel critical assembly, a team at ORNL preformed an investigation of a hypothetical transportation package with 20% enriched 235U in the form of UF6 using ENDF/B-VII.1 and ENDF/B-VIII.0 libraries by only exchanging the CIELO nuclides. A swing in $k_{eff}$ of 450 pcm is observed.  Future investigation and validation are required to gain confidence in either calculation.

@@

Following NDNCA, Dr. Catherine Romano, working with the DOE-NP and the NNSA office on counter-proliferation research and development (NA-22), organized a grassroots \enquote{Nuclear Data Working Group} (NDWG) comprised of subject matter experts selected by program managers from DOE-NP, DOE-Nuclear Energy, the Isotopes Program, the Department of Homeland Security and the several offices within the National Nuclear Security Administration to identify a \enquote{roadmap} to address the most important and crosscutting nuclear data needs including all steps in the nuclear data pipeline from measurement through compilation, evaluation and processing.  The NDWG presented their results to participating program managers as well as other interested agencies during the Nuclear Data Exchange Meeting (NDEM) in April 2016.  The scope of the recommended work included:
1. infrastructure modernization, 
2. expanded covariance data, 
3. inelastic scatter on 235U, 238U and 239Pu, 
4. gamma emission data, 
5. fission yields, and 
6. renewed nuclear data target production capabilities.
The NDEM provided a venue for valuable discussions between the nuclear data community and program managers and the tasks required to provide evaluated nuclear data to the database.
